How accurate is CC4 for vertical excitation energies?

### 2021-05-01 Manuscript/CC4.tex
writing intro

\documentclass[aps,prb,reprint,noshowkeys,superscriptaddress]{revtex4-1}
\usepackage{graphicx,dcolumn,bm,xcolor,microtype,multirow,amscd,amsmath,amssymb,amsfonts,physics,wrapfig,txfonts}
\usepackage[version=4]{mhchem}

\newcommand{\ie}{\textit{i.e.}}
\newcommand{\eg}{\textit{e.g.}}
\newcommand{\alert}[1]{\textcolor{black}{#1}}
\usepackage[normalem]{ulem}
\newcommand{\titou}[1]{\textcolor{red}{#1}}
\newcommand{\trashPFL}[1]{\textcolor{red}{\sout{#1}}}
\newcommand{\PFL}[1]{\titou{(\underline{\bf PFL}: #1)}}
\newcommand{\toto}[1]{\textcolor{green}{#1}}
\newcommand{\trashAS}[1]{\textcolor{green}{\sout{#1}}}
\newcommand{\AS}[1]{\toto{(\underline{\bf AS}: #1)}}

\newcommand{\Ec}{E_\text{c}}
\newcommand{\mEh}{$mE_h$}
\newcommand{\Eh}{$E_h$}
\newcommand{\mc}{\multicolumn}
\newcommand{\fnm}{\footnotemark}
\newcommand{\fnt}{\footnotetext}
\newcommand{\tabc}[1]{\multicolumn{1}{c}{#1}}
\newcommand{\QP}{\textsc{quantum package}}

@@

\newcommand{\LCPQ}{Laboratoire de Chimie et Physique Quantiques (UMR 5626), Universit\'e de Toulouse, CNRS, UPS, France}
\newcommand{\CEISAM}{Universit\'e de Nantes, CNRS,  CEISAM UMR 6230, F-44000 Nantes, France}

@@

\author{Pierre-Fran\c{c}ois Loos}
\email{loos@irsamc.ups-tlse.fr}
\affiliation{\LCPQ}
\author{Denis Jacquemin}
\email{Denis.Jacquemin@univ-nantes.fr}
\affiliation{\CEISAM}

% Abstract
\begin{abstract}
\end{abstract}

@@

@@

@@

\begin{squeezetable}
\begin{table*}
	\caption{Vertical excitation energies (in eV) of a selection of molecular excited states obtained at various levels of theory with the aug-cc-pVDZ and aug-cc-pVTZ basis sets.
	\label{tab:BigTab}}
	\begin{ruledtabular}
	\begin{tabular}{llrrrrrrrrrrrr}
				&		&	\mc{7}{c}{aug-cc-pVDZ}		&		\mc{5}{c}{aug-cc-pVTZ}		\\	
				\cline{3-9} \cline{10-14}
	Mol.	&	State				&CC3	&CCSDT-3&CCSDT	&CC4	&CCSDTQ	&CCSDTQP	&CIPSI	
									&CC3	&CCSDT	&CC4	&CCSDTQ	&CIPSI			\\
	\hline
	\ce{NH3}	&	$^1A_2$ 		&6.464	&6.472	&6.462	&6.479	&6.480	&6.482	&6.483(1)	&6.573	&6.571	&6.585	&6.586	&	6.593(22)	\\	
				&	$^1E$			&8.061	&8.065	&8.057	&8.078	&8.079	&8.081	&8.082(1)	&8.146	&8.143	&8.161	&8.161	&	8.171(20)	\\	
				&	$^1A_1$ 		&9.664	&9.668	&9.659	&9.677	&9.677	&9.680	&9.681(8)	&9.318	&9.314	&9.331	&9.331	&	9.340(19)	\\
				&	$^1A_2$ 		&10.396	&10.402	&10.391	&10.409	&10.409	&10.411	&10.412(1)	&9.945	&9.939	&9.957	&9.957	&	9.967(19)	\\
	\ce{BH}		&	$^1\Pi$ 		&2.955	&2.958	&2.946	&2.947	&2.947	&2.947	&2.947(0)	&2.910	&2.900	&2.901	&2.901	&2.901(0)	\\
	\ce{BF}		&	$^1\Pi$ 		&6.478	&6.491	&6.491	&6.484	&6.485	&6.485	&6.485(1)	&6.410	&6.423	&6.416	&6.417	&6.418(2)\\
	\ce{CO}		&	$^1\Pi$ 		&8.572	&8.591	&8.574	&8.562	&8.563	&8.561	&8.565(2)	&8.486	&8.492	&8.479	&8.480	&	\\
				&	$^1\Sigma^-$ 	&10.122	&10.112	&10.062	&10.055	&10.057	&10.057	&10.056(1)	&9.992	&9.940	&9.930	&9.932	&	\\
				&	$^1\Delta$ 		&10.225	&10.220	&10.178	&10.167	&10.169	&10.168	&10.168(1)	&10.119	&10.076	&10.064	&10.066	&	\\
				&	$^1\Sigma^+$ 	&10.917	&10.989	&10.944	&10.925	&10.926	&10.919	&			&10.943	&10.987	&10.961	&10.963	&	\\
				&	$^1\Sigma^+$ 	&11.483	&11.543	&11.518	&11.510	&11.510	&11.506	&			&11.489	&11.540	&11.521	&11.523	&	\\
				&	$^1\Pi$ 		&11.737	&11.802	&11.767	&11.757	&11.758	&11.753	&			&11.690	&11.737	&11.719	&11.720	&	\\
	\ce{N2}		&	$^1\Pi_g$  		&9.442	&9.450	&9.417	&9.409	&9.411	&9.409	&9.411(3)	&9.344	&9.326	&9.317	&9.319	&	\\
				&	$^1\Sigma_u^-$	&10.059	&10.063	&10.060	&10.063	&10.055	&10.054	&10.054(0)	&9.885	&9.890	&9.883	&9.878	&	\\
				&	$^1\Delta_u$ 	&10.433	&10.449	&10.436	&10.439	&10.430	&10.428	&10.429(0)	&10.293	&10.302	&10.294	&10.287	&	\\
				&	$^1\Sigma_g^+$	&13.229	&13.270	&13.202	&13.171	&13.182	&13.181	&13.180(1)	&13.013	&12.999	&12.962	&12.974	&	\\
				&	$^1\Pi_u$  		&13.279	&13.323	&13.174	&13.128	&13.131	&13.127	&			&13.223	&13.140	&13.091	&13,095	&	\\
				&	$^1\Sigma_u^+$	&13.146	&13.186	&13.130	&13.099	&13.109	&13.107	&			&13.120	&13.118	&13.078	&13.090	&	\\
				&	$^1\Pi_u$ 		&13.635	&13.663	&13.591	&13.551	&13.560	&13.558	&			&13.494	&13.455	&13.409	&13.419\\
	\ce{HCl}	&	$^1\Pi$ 		&7.819	&7.828	&7.815	&7.822	&7.822	&7.823	&7.823(0)	&7.840	&7.834	&7.837	&7.837	&7.838(1)\\
	\ce{H2S}	&	$^1B_1$  		&6.098	&6.110	&6.098	&6.102	&6.103	&6.103	&6.103(1)	&6.240	&6.237	&6.238	&6.238	&6.240(7)\\
				&	$^1A_2$  		&6.293	&6.299	&6.286	&6.286	&6.286	&6.286	&6.286(0)	&6.192	&6.185	&6.181	&6.181	&6.181(6) \\
	\ce{H2O}	&	$^1B_1$  		&7.511	&7.510	&7.497	&7.531	&7.528	&7.532	&7.533(0)	&7.605	&7.591	&7.623	&7.620	&7.626(3)\\
				&	$^1A_2$  		&9.293	&9.291	&9.279	&9.317	&9.313	&9.318	&9.318(0)	&9.382	&9.368	&9.405	&9.400	&9.407(7)\\
				&	$^1A_1$  		&9.921	&9.919	&9.903	&9.940	&9.937	&9.941	&9.941(0)	&9.966	&9.949	&9.986	&9.981	&9.987(8)\\
	\hline
	MAE			&					&	\\
	\end{tabular}
	\end{ruledtabular}
\end{table*}
\end{squeezetable}

@@

@@

\section*{Data availability statement}
The data that support the findings of this study are openly available in Zenodo at \href{http://doi.org/XX.XXXX/zenodo.XXXXXXX}{http://doi.org/XX.XXXX/zenodo.XXXXXXX}.
\bibliography{Ec}

>>>>

\documentclass[aip,jcp,reprint,noshowkeys,superscriptaddress]{revtex4-1}
\usepackage{graphicx,dcolumn,bm,xcolor,microtype,multirow,amscd,amsmath,amssymb,amsfonts,physics,wrapfig,txfonts,siunitx}
\usepackage[version=4]{mhchem}

\newcommand{\alert}[1]{\textcolor{black}{#1}}
\usepackage[normalem]{ulem}
\newcommand{\titou}[1]{\textcolor{red}{#1}}
\newcommand{\trashPFL}[1]{\textcolor{red}{\sout{#1}}}
\newcommand{\PFL}[1]{\titou{(\underline{\bf PFL}: #1)}}

\newcommand{\ie}{\textit{i.e.}}
\newcommand{\eg}{\textit{e.g.}}
\newcommand{\mc}{\multicolumn}
\newcommand{\fnm}{\footnotemark}
\newcommand{\fnt}{\footnotetext}
\newcommand{\tabc}[1]{\multicolumn{1}{c}{#1}}

% Softwares
\newcommand{\QP}{\textsc{quantum package}}
\newcommand{\MRCC}{\textsc{mrcc}}
\newcommand{\CFOUR}{\textsc{cfour}}

% variables
\newcommand{\hH}{\Hat{H}}
\newcommand{\hT}{\Hat{T}}
\newcommand{\bH}{\Bar{H}}

@@

\newcommand{\LCPQ}{Laboratoire de Chimie et Physique Quantiques (UMR 5626), Universit\'e de Toulouse, CNRS, UPS, France}
\newcommand{\SMU}{Department of Chemistry, Southern Methodist University, Dallas, Texas 75275, USA}
\newcommand{\UOP}{Dipartimento di Chimica e Chimica Industriale, University of Pisa, Via Moruzzi 3, 56124 Pisa, Italy}
\newcommand{\CEISAM}{Universit\'e de Nantes, CNRS,  CEISAM UMR 6230, F-44000 Nantes, France}

@@

\author{Pierre-Fran\c{c}ois Loos}
	\email{loos@irsamc.ups-tlse.fr}
	\affiliation{\LCPQ}
\author{Devin A.~Matthews}
	\affiliation{\SMU}
\author{Filippo Lipparini}
	\affiliation{\UOP}
\author{Denis Jacquemin}
	\email{Denis.Jacquemin@univ-nantes.fr}
	\affiliation{\CEISAM}

% Abstract
\begin{abstract}
\titou{Here comes the abstract.}
\end{abstract}

@@

The hierarchy of coupled cluster (CC) methods provide a route to systematically improvable energies and properties via the systematic increase of the maximum excitation degree of the cluster operator $\hT = \hT_1 + \hT_2 + \ldots + \hT_N$ (where $N$ is the number of electrons).
Without any truncation, full CC (FCC) is equivalent to full configuration interaction (FCI) and provides the exact energy of the system. 
However, it is not computationally viable and one usually resorts to truncated CC methods.
Popular choices are CC with singles and doubles (CCSD), CC with singles, doubles, and triples (CCSDT), CC with singles, doubles, triples, and quadruples (CCSDTQ), and 
CC with singles, doubles, triples, quadruples, and pentuples (CCSDTQP) with corresponding computational scalings of $\order*{M^{6}}$, $\order*{M^{8}}$,  $\order*{M^{10}}$, and  $\order*{M^{12}}$, respectively (where $M$ denotes the number of orbitals).
An alternative systematically-improvable family of methods is provided by the CC2, CC3, and CC4 models which have been introduced by the Aarhus group in the context of CC response theory.
These methods scale as $\order*{M^{5}}$, $\order*{M^{7}}$, and $\order*{M^{9}}$, respectively, and can be seen as a cheap approximation of CCSD, CCSDT, and CCSDTQ, by skipping the most expensive terms and avoiding the storage of the higher-excitation amplitudes.
A similar strategy has been applied to produce the CCSDT-3 and CCSDTQ-3 models based on arguments stemming from perturbation theory.
Of course, many more approximate CC methods have been developed and we refer the interested reader to specialized reviews for more details.

Coupled-cluster methods have been particularly successful for small- and medium-sized molecules in the field of thermodynamics and spectroscopy, thanks to the computations of accurate geometries, potential energy surfaces, couplings between electronic states, a vast panel of properties such as oscillator strengths as well as ground- and excited-state dipoles.
Although originally developed from ground-state energies and properties, CC has been successfully extended to excited states thanks to the equation-of-motion (EOM) and linear-response (LR) formalisms which are known to produce identical excitation energies but different properties.
In EOM-CC, one obtains vertical excitation energies via the diagonalization of the similarity-transformed Hamiltonian $\bH = e^{-\hT} \hH e^{\hT}$.
A more general procedure to compute excitation energies that can be straightforwardly applied to any truncated CC family as the ones mentioned above consists in diagonalizing the so-called CC Jacobian obtained via the differentiation of the CC amplitude equations.

The purpose of the present study is to assess the accuracy of the approximate CC4 model against the most expensive CCSDTQ and CCSDTQP methods.
near-FCI results.

@@

Each method used in the present study is reported in Table \ref{tab:scaling} alongside its formal computational scaling and the electronic structure software used to compute the excitation energies.

@@

\begin{table}
	\caption{Methods considered, their formal computational scaling and the name of the electronic structure software where it has been implemented.
	Here, $M$ is the number of orbitals.
	\label{tab:scaling}}
	\begin{ruledtabular}
	\begin{tabular}{lccc}
		Method	&	Scaling					&	Code		&	Ref.					\\
		\hline
		CC3			&	$\order*{M^{7}}$	&	\CFOUR		&	\onlinecite{cfour}		\\
		CCSDT-3		&	$\order*{M^{7}}$	&	\CFOUR		&	\onlinecite{cfour}		\\
		CCSDT		&	$\order*{M^{8}}$	&	\CFOUR		&	\onlinecite{cfour}		\\
		CC4			&	$\order*{M^{9}}$	&	\CFOUR		&	\onlinecite{cfour}		\\
		CCSDTQ-3	&	$\order*{M^{9}}$	&	\CFOUR		&	\onlinecite{cfour}		\\
		CCSDTQ		&	$\order*{M^{10}}$	&	\CFOUR		&	\onlinecite{cfour}		\\
		CCSDTQP		&	$\order*{M^{12}}$	&	\MRCC		&	\onlinecite{mrcc}		\\	
		CIPSI		&	$\order*{e^{M}}$	&	\QP			&	\onlinecite{qp2}		\\
	\end{tabular}
	\end{ruledtabular}
\end{table}

\begin{squeezetable}
\begin{table*}
	\caption{Vertical excitation energies (in eV) of a selection of molecular excited states obtained at various levels of theory with the aug-cc-pVDZ and aug-cc-pVTZ basis sets.
	\label{tab:BigTab}}
	\begin{ruledtabular}
	\begin{tabular}{llrrrrrrrrrrrr}
				&		&	\mc{7}{c}{aug-cc-pVDZ}		&		\mc{5}{c}{aug-cc-pVTZ}		\\	
				\cline{3-9} \cline{10-14}
	Mol.	&	State				&CC3	&CCSDT-3&CCSDT	&CC4	&CCSDTQ	&CCSDTQP	&CIPSI	
									&CC3	&CCSDT	&CC4	&CCSDTQ	&CIPSI			\\
	\hline
	\ce{NH3}	&	$^1A_2$ 		&6.464	&6.472	&6.462	&6.479	&6.480	&6.482	&6.483(1)	&6.573	&6.571	&6.585	&6.586	&	6.593(22)	\\	
				&	$^1E$			&8.061	&8.065	&8.057	&8.078	&8.079	&8.081	&8.082(1)	&8.146	&8.143	&8.161	&8.161	&	8.171(20)	\\	
				&	$^1A_1$ 		&9.664	&9.668	&9.659	&9.677	&9.677	&9.680	&9.681(8)	&9.318	&9.314	&9.331	&9.331	&	9.340(19)	\\
				&	$^1A_2$ 		&10.396	&10.402	&10.391	&10.409	&10.409	&10.411	&10.412(1)	&9.945	&9.939	&9.957	&9.957	&	9.967(19)	\\
	\ce{BH}		&	$^1\Pi$ 		&2.955	&2.958	&2.946	&2.947	&2.947	&2.947	&2.947(0)	&2.910	&2.900	&2.901	&2.901	&2.901(0)	\\
	\ce{BF}		&	$^1\Pi$ 		&6.478	&6.491	&6.491	&6.484	&6.485	&6.485	&6.485(1)	&6.410	&6.423	&6.416	&6.417	&6.418(2)\\
	\ce{CO}		&	$^1\Pi$ 		&8.572	&8.591	&8.574	&8.562	&8.563	&8.561	&8.565(2)	&8.486	&8.492	&8.479	&8.480	&	\\
				&	$^1\Sigma^-$ 	&10.122	&10.112	&10.062	&10.055	&10.057	&10.057	&10.056(1)	&9.992	&9.940	&9.930	&9.932	&	\\
				&	$^1\Delta$ 		&10.225	&10.220	&10.178	&10.167	&10.169	&10.168	&10.168(1)	&10.119	&10.076	&10.064	&10.066	&	\\
				&	$^1\Sigma^+$ 	&10.917	&10.989	&10.944	&10.925	&10.926	&10.919	&			&10.943	&10.987	&10.961	&10.963	&	\\
				&	$^1\Sigma^+$ 	&11.483	&11.543	&11.518	&11.510	&11.510	&11.506	&			&11.489	&11.540	&11.521	&11.523	&	\\
				&	$^1\Pi$ 		&11.737	&11.802	&11.767	&11.757	&11.758	&11.753	&			&11.690	&11.737	&11.719	&11.720	&	\\
	\ce{N2}		&	$^1\Pi_g$  		&9.442	&9.450	&9.417	&9.409	&9.411	&9.409	&9.411(3)	&9.344	&9.326	&9.317	&9.319	&	\\
				&	$^1\Sigma_u^-$	&10.059	&10.063	&10.060	&10.063	&10.055	&10.054	&10.054(0)	&9.885	&9.890	&9.883	&9.878	&	9.879(4)\\
				&	$^1\Delta_u$ 	&10.433	&10.449	&10.436	&10.439	&10.430	&10.428	&10.429(0)	&10.293	&10.302	&10.294	&10.287	&	10.289(12)\\
				&	$^1\Sigma_g^+$	&13.229	&13.270	&13.202	&13.171	&13.182	&13.181	&13.180(1)	&13.013	&12.999	&12.962	&12.974	&	\\
				&	$^1\Pi_u$  		&13.279	&13.323	&13.174	&13.128	&13.131	&13.127	&			&13.223	&13.140	&13.091	&13,095	&	\\
				&	$^1\Sigma_u^+$	&13.146	&13.186	&13.130	&13.099	&13.109	&13.107	&			&13.120	&13.118	&13.078	&13.090	&	\\
				&	$^1\Pi_u$ 		&13.635	&13.663	&13.591	&13.551	&13.560	&13.558	&			&13.494	&13.455	&13.409	&13.419\\
	\ce{HCl}	&	$^1\Pi$ 		&7.819	&7.828	&7.815	&7.822	&7.822	&7.823	&7.823(0)	&7.840	&7.834	&7.837	&7.837	&7.838(1)\\
	\ce{H2S}	&	$^1B_1$  		&6.098	&6.110	&6.098	&6.102	&6.103	&6.103	&6.103(1)	&6.240	&6.237	&6.238	&6.238	&6.240(7)\\
				&	$^1A_2$  		&6.293	&6.299	&6.286	&6.286	&6.286	&6.286	&6.286(0)	&6.192	&6.185	&6.181	&6.181	&6.181(6) \\
	\ce{H2O}	&	$^1B_1$  		&7.511	&7.510	&7.497	&7.531	&7.528	&7.532	&7.533(0)	&7.605	&7.591	&7.623	&7.620	&7.626(3)\\
				&	$^1A_2$  		&9.293	&9.291	&9.279	&9.317	&9.313	&9.318	&9.318(0)	&9.382	&9.368	&9.405	&9.400	&9.407(7)\\
				&	$^1A_1$  		&9.921	&9.919	&9.903	&9.940	&9.937	&9.941	&9.941(0)	&9.966	&9.949	&9.986	&9.981	&9.986(2)\\
	\hline
	MAE			&					&	\\
	\end{tabular}
	\end{ruledtabular}
\end{table*}
\end{squeezetable}

@@

\titou{Here comes the conclusion.}

@@

\section*{Data availability statement}
The data that support the findings of this study are openly available in Zenodo at \href{http://doi.org/XX.XXXX/zenodo.XXXXXXX}{http://doi.org/XX.XXXX/zenodo.XXXXXXX}.

\bibliography{CC4}
